\documentclass[letterpaper,12pt,titlepage]{article}

\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}

\usepackage{alltt}
\usepackage{float}
\usepackage{color}
\usepackage{url}

\usepackage[framemethod=TikZ]{mdframed}
\mdfdefinestyle{MyFrame}{%
    innertopmargin=\baselineskip,
    innerbottommargin=\baselineskip,
    innerrightmargin=20pt,
    innerleftmargin=20pt,
    backgroundcolor=gray!5!white,
    splitbottomskip = 5mm,
    splittopskip = 5mm,
    skipabove=5mm}

\usepackage{listings}
\usepackage{xcolor}
\usepackage{color}
\usepackage[font=small,format=plain,labelfont=bf,up,textfont=it,up]{caption}

\usepackage{balance}
\usepackage[TABBOTCAP, tight]{subfigure}
\usepackage{enumitem}
\usepackage{pstricks, pst-node}

%%keeps text of subsection on same line
\usepackage{titlesec}
\titleformat{\subsection}[runin]
  {\normalfont\large\bfseries}{\thesubsection}{1em}{}

\usepackage{geometry}
\geometry{textheight=9in, textwidth=6.5in}

%random comment

\newcommand{\cred}[1]{{\color{red}#1}}
\newcommand{\cblue}[1]{{\color{blue}#1}}

\usepackage{hyperref}
\usepackage{geometry}

%% Code Listing Configuration
\renewcommand{\lstlistingname}{Code}
\lstset{ %
  %Some lang opts: C++, C, Java, Python, Matlab, TeX, HTML, SQL, Verilog, VHDL, make, ...
  basicstyle=\footnotesize\ttfamily , % the size of the fonts that are used for the code
  numbers=left,                       % where to put the line-numbers
  numberstyle=\scriptsize\color{darkgray}, % the style that is used for the line-numbers
  stepnumber=2,                       % the step between two line-numbers.
  numbersep=5pt,                      % how far the line-numbers are from the code
  backgroundcolor=\color{white},      % choose the background color. You must add \usepackage{color}
  showspaces=false,                   % show spaces adding particular underscores
  showstringspaces=false,             % underline spaces within strings
  showtabs=false,                     % show tabs within strings adding particular underscores
  frame=tb,                           % adds a frame around the code
  rulesepcolor=\color{gray},          % if not set, the frame-color may be changed on line-breaks within not-black text (e.g. commens (green here))
  tabsize=2,                          % sets default tabsize to 2 spaces
  captionpos=t,                       % sets the caption-position
  breaklines=true,                    % sets automatic line breaking
  breakatwhitespace=false,            % sets if automatic breaks should only happen at whitespace
  title=\lstname,                     % show the filename of files included with \lstinputlisting;
  keywordstyle=\color{blue},          % keyword style
  commentstyle=\color{dkgreen},       % comment style
  stringstyle=\color{mauve},          % string literal style
  escapeinside={\%*}{*)},             % if you want to add a comment within your code
  morekeywords={*,...}                % if you want to add more keywords to the set
  framexbottommargin=5pt,
}

\def\name{Drake Bridgewater \& Ryan Phillips}

%% The following metadata will show up in the PDF properties
\hypersetup{
  colorlinks = true,
  urlcolor = black,
  pdfauthor = {\name},
  pdfkeywords = {cs472 ``computer architecture'' clements ``chapter 3''},
  pdftitle = {CS 472: Homework 4},
  pdfsubject = {CS 472: Homework 4},
  pdfpagemode = UseNone
}

\begin{document}
\hfill \name

\hfill \today

\hfill CS 472 HW 4

\section*{$(3.8)$} What are the relative advantages and disadvantages of general-purpose registers compared to separate address and data registers?

\begin{mdframed}[style=MyFrame]
\end{mdframed}

\section*{$(3.9)$} What is a misaligned operand? Why are misaligned operands such a problem in programming?

\begin{mdframed}[style=MyFrame]
A misaligned operand is just what it sounds like an operand that is not aligned causing the operand to span two cache lines either slowing the system down substantially or causing a bus error. From some articles online typically x86 and x64 will happily work with unaligned data but will tend to be slower where as on other processors the system will just error. From the ARM website "Non word-aligned load and store multiple, double, semaphore, synchronization, and coprocessor accesses always signal Data Abort with an Alignment fault status code when the U bit is set." \cite{ARM}
\end{mdframed}

\section*{$(3.24)$} What is the meaning of each of the P,U,B,W, and L bits in the encoding of an ARM memory reference instruction?
\begin{mdframed}[style=MyFrame]
\begin{tabular}{l l}
\textbf{P} & Do you want to adjust pointer before using?\\
\textbf{U} & Do you want to decrement the pointer instead of incrementing it?\\
\textbf{B} & Is this word access?\\
\textbf{W} & Do you want to update the pointer after use?\\
\textbf{L} & Do you want to store data in memory?
\end{tabular}
\end{mdframed}

\section*{$(3.26)$} What is the effect of LDR r0, [r5, r6, LSL r2] ?

\begin{mdframed}[style=MyFrame]
<<<<<<< HEAD
\colorbox{red}{
  
Whatever is at the address located specified by: address r5 + (address r6)*2^r2 is loaded into r0.


}
=======
$[r0] \leftarrow [r5+[r6 * r2] ]$ or in english store what is in R5 with the addition of the offset determined by logically shifting r6 left by r2 in r0
>>>>>>> a70a315180d74f562ef1f5b38601446208480cb4
\end{mdframed}

\section*{$(3.30)$} What is the meaning of sign-extension in the context of copying data from one location to another?

\begin{mdframed}[style=MyFrame]
Sign extension is maintaning the binary sign when changing the length of represented value. For instance if you are moving a 16-bit value into a 32-bit register you have to ensure that you sign extend the bits this means that the most significant bit will need to be repeated until the end in the new register.  
\end{mdframed}

\section*{$(3.33)$} Most RISC processors do not include a block move instruction. What are the advantages and disadvantages of the ARM's LDM and STM instructions?

\begin{mdframed}[style=MyFrame]
Using a block load or store instruction can be excuted in few cycles for more then a single register. For instance if you want to block move 16 register it will take 8 cycles if 
\end{mdframed}

\newpage
\section*{$(3.34)$} What is the effect of executing STMIB r13!,\{r0-r2,r4\}? Draw a picture of the state of the stack pointed at by r13 before and after this operation.
\begin{mdframed}[style=MyFrame]

The instruction STMIB is the Store Multiple Registers instruction and it increments the address before each transfer and with the addition of the exclimation mark on Rd the final address is written back to Rd, R13 \\

\begin{center}
Assuming this is a Full Ascending, decrement after stack
\begin{tabular}{| c | c | c |}
\hline
& Before & After\\ \hline \hline
\& $n+0$ &  $\rightarrow$ &$~~~~R0$\\ \hline
\& $n+4$ &   &$~~~~R1$\\ \hline
\& $n+8$ &   &$~~~~R2$\\ \hline
\& $n+C$ &   &$\rightarrow$ $R4$\\ \hline
\& $n+F$ &   &Free\\ \hline

\end{tabular}
\end{center}
\end{mdframed}

\section*{$(3.36)$} Without using the ARM's multiplication instruction, write one or more instructions (using ADD, SUB, and shifting) to multiply by the following integers: 33, 1025, 4095

\begin{mdframed}[style=MyFrame]
In binary: \\
\begin{tabular}{l l }
33 & 0010 0001 \\
1025 & 0100 0000 0001 \\
4095 & 1111 1111 1111 \\ \\
$33\times 1025 = 33825$ & 1000 0100 0010 0001
\end{tabular}

\begin{verbatim}
    AREA test, CODE, READWRITE    
        ENTRY
        
        ; multiply input by 33 -> original += 4 byte shift
        MOV        R0, #4095              ; 
        MOV        R1, R0, LSL #5         ; r1 = r0 * 2^5 
        ADD        R2, R1, R0             ; r2 = r1 + r0
                                          ; r2 contains r0 * 33
        space     10                            
        ; multiply input by 33 -> original += 4 byte shift
        MOV        R0, R2                 ; 
        MOV        R1, R0, LSL #10        ; r1 = r0 * 2^5 
        ADD        R2, R1, R0             ; r2 = r1 + r0
        
        space     10
    
        END
\end{verbatim}
\end{mdframed}

\section*{$(3.44)$} What does the following code do?
\begin{verbatim}
TEQ     r0, #0
RSBMI   r0, r0, #0
\end{verbatim}

\begin{mdframed}[style=MyFrame]
\textbf{TEQ} is the instruction test equivalence it uses a bitwise exclusive OR operation when it is complete it sets the flags and the result is discarded unlike the EORS instruction\\
\textbf{RSBMI} is the instruction Reverse Subtract without carry but the addition of MI means that it will only execute when N status register is set, it is negative. 

TEQ will set the N, negative flag, for any negative value. If the value is a negative then RSBMI instruction will execute; yielding 0 - R0. Overall these two instructions used together will take the absolute value of R0.
\end{mdframed}

\section*{$(3.48)$} What, in the context of assembly language, is a psuedo-operation?

\begin{mdframed}[style=MyFrame]
These are keywords which do not directly translate to a machine instructions
\end{mdframed}
\newpage
\section*{$(3.54)$} Explain what this fragment of code does instruction by instruction and what purpose it achieves (assuming that register r0 is the register of interest). Note that the data in r0 must not be 0 on entry.
\begin{verbatim}
        MOV        r1,#0
loop    MOVS       r0,r0,LSL #1
        ADDCC      r1,r1,#1
        BCC        loop
\end{verbatim}

\begin{mdframed}[style=MyFrame]
Moves 0 into R1\\
Moves R0 left shifted by 1 into R0 and updated the status register\\
Adds $R1 = R1 + 1$ if carry then flag is set\\
Loops back to loop if carry then flag is set\\ \\

This will multiply the value in R0 by 2 each until the most significant bit is a 1 and the value accrued in R1 is the value that is was multiplied by. 
\end{mdframed}

\newpage
\section*{$(3.60)$} A computer has three eight-element vectors in memory, Va, Vb, and Vc. Each element of a vector is a 32-bit word. Write the code to calculate all elements of Vc if the ith element is given by: $Vc_i = \frac{1}{2} ( Va_i + Vb_i)$
\begin{mdframed}[style=MyFrame]
\lstinputlisting[linerange=1-40]{../Lab3/quad_func/quad_func.asm}
\end{mdframed}
\newpage

\section*{$(3.61)$} Register r15 is the program counter. You can use it with certain instructions such as a MOV (e.g., MOV pc, r14). However, r15 cannot be used in conjunction with most data processing instructions. Why?

\begin{mdframed}[style=MyFrame]
Since the program counter is not connected to the ALU instuctions can't be directly executed using the value in the program counter. If you want to use the program counter you must load it into a register then manipulate that register
\end{mdframed}

\bibliographystyle{plain}
\bibliography{writeup}
\end{document}
